Automated Program Verification and Synthesis of Embedded \& Cyber-physical Systems: Advances and Challenges

### 2017-08-16 paper.tex
fixed typos
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

The overall automated formal synthesis process for ECPSs and HSs involves three main steps: \textit{modeling}, which is related to obtaining a sound mathematical representation able to capture all relevant aspects of continuous and discrete dynamics and their interaction, as well as formulating properties and invariants that describe a desirable behavior, \textit{verification}, where a ECPS's behavior is checked through the proposed mathematical structure and compliance with a desired behavior is verified, and, finally, controller \textit{synthesis}, which produces a controller or finds invariant control laws that allow a ECPS to meet requirements and behave as desired, which is confirmed by a verification oracle.

@@

Another approach to ECPS modeling tackles quantized control systems and finite-word length (FWL) effects on the performance of those systems. Differently from bisimulation, quantized models are built from a countable subset of the (quantized) input space~\cite{Tabuada07}. In particular, Keel and Bhattacharyya showed that even robust and optimal controllers may still present an undesired property~\cite{Keel01,bhattacharyya}: fragility or sensitivity to implementation aspects ({\it e.g.}, FWL representations of digital controllers). Therefore, various studies tried link sampled-data theory to hybrid-system models~\cite{Bicchi02,Petreczky10,Petreczky07,Ye98}.

@@

Given a class of ECPSs $\Omega$ and a set $\Phi$ of desired properties, a class of verification problems is called decidable if there is an algorithm that is able to decide if a finite number of steps $\forall \mathcal{O}\in\Omega$ satisfy every property $\mathcal{L}\in \Phi$~\cite{Alur00}. Decidability is a key issue in ECPS verification, since there are uncountable states and complex dynamics to be considered, {\it i.e.}, hybrid automata and their variations (usually employed to model ECPSs) are infinite-state machines~\cite{Henzinger95}. In particular, the safety verification of ECPS is undecidable, except in very specific cases as in systems that can be represented by timed automata and rectangular automata~\cite{Alur11}. An alternative to ensure that ECPS trajectories will not pass through unsafe states is the use of barrier certificates~\cite{Prajna07,Prajna04}. An example of barrier certificate is the stability, that ensures the convergence of states to an equilibrium point and consequently that there exists a region from where any trajectory remains inside of this region. The stability may be ensured based on Lyapunov theory (for a wide class of model abstractions)~\cite{tabuada2009verification}, or linear system stability theory for discrete simulations~\cite{Bessa17}.

Generally, properties that must hold for ECPSs are related to safety/reachability and liveness, and the earliest works on HS and ECPS proposed approaches based on computational tree logic and linear temporal logic~\cite{Alur93}, in order to specify properties and verify those systems. In the last decade, the signal temporal logic~\cite{Maler04,Donze12} was proposed and have been used since then for this purpose, thus providing more suitable operators for dynamic systems and signal processing. Therefore, model-checking procedures have been employed to verify those properties, even under undecidability conditions via approximations, {\it e.g.}, bounding the number of steps from initial states, which leads to bounded model checking (BMC)~\cite{Veanes09}.

@@

Finally, it is worth mentioning that there are other alternatives for ECPS modeling and analysis, which are not based on verification, but instead on simulators~\cite{Phan2014,Nakajima12,Junjie12}. Although such approaches generally fail in ensuring that some properties hold, they present good scalability, which is ideal for large scale and complex systems that are hardly modeled through automata- or logig-based approaches. Modelica~\cite{Junjie12} and Simulink/MATLAB~\cite{Nakajima12} are important examples of widely used simulators for ECPSs. In order to achieve a balanced trade-off between ECPS-simulation scalability and correctness of verification methods, statistical model checking have been employed to formally verify ECPSs~\cite{Clarke11}.

In addition, Shoukry {\it et al.} \cite{Shoukry} tackled another important problem related to ECPSs: security. Indeed, the essence of ECPSs claims for sophisticated interfacing ({\it e.g.}, sensors and network connections), which normally leads to security issues~\cite{cpsattack}, given that wrong control actions can be generated through corrupted measurements. The mentioned authors proposed a methodology for estimating physical-system state, through formal methods, even if sensor inputs are contaminated with erroneous or intentionally corrupted data, in order to still be able to execute suitable control commands. Cyber-security is also a concern of Fiore {\it et al.}, who aim to ensure secure-state estimation of UAV systems under sparse attacks~\cite{Fiore17}. Indeed, many situations can be regarded as satisfiability problems and solved, for instance, with approaches based on Satisfiability Modulo Theories (SMT)~\cite{Araujo16,TrindadeDAES2016,Rahman13}, which indicates many links with other research areas and might lead to a unified approach regarding system verification, security enforcement, and interaction handling.

%--------------------------------------
\subsection{Control synthesis of Cyber-physical systems}
\label{ssec:synthesis}
%--------------------------------------

>>>>

The overall automated formal synthesis process for ECPSs and HSs involves three main steps: \textit{modeling}, which is related to obtaining a sound mathematical representation able to capture all relevant aspects of continuous and discrete dynamics and their interaction, as well as formulating properties and invariants that describe a desirable behavior, \textit{verification}, where a ECPS behavior is checked through the proposed mathematical structure and compliance with a desired behavior is verified, and, finally, controller \textit{synthesis}, which produces a controller or finds invariant control laws that allow a ECPS to meet requirements and behave as desired, which is confirmed by a verification oracle.

@@

Another approach to ECPS modeling tackles quantized control systems and finite-word length (FWL) effects on the performance of those systems. Differently from bisimulation, quantized models are built from a countable subset of the (quantized) input space~\cite{Tabuada07}. In particular, Keel and Bhattacharyya showed that even robust and optimal controllers may still present an undesired property~\cite{Keel01,bhattacharyya}: fragility or sensitivity to implementation aspects ({\it e.g.}, FWL representations of digital controllers). Therefore, various studies tried to connect sampled-data theory to hybrid-system models~\cite{Bicchi02,Petreczky10,Petreczky07,Ye98}.

@@

Given a class of ECPSs $\Omega$ and a set $\Phi$ of desired properties, a class of verification problems is called decidable if there is an algorithm that is able to decide if a finite number of steps $\forall \mathcal{O}\in\Omega$ satisfy every property $\mathcal{L}\in \Phi$~\cite{Alur00}. Decidability is a key issue in ECPS verification, since there are uncountable states and complex dynamics to be considered, {\it i.e.}, hybrid automata and their variations (usually employed to model ECPSs) are infinite-state machines~\cite{Henzinger95}. In particular, the safety verification of ECPS is undecidable, except in very specific cases as in systems that can be represented by timed automata and rectangular automata~\cite{Alur11}. An alternative to ensure that ECPS trajectories will not pass through unsafe states is the use of barrier certificates~\cite{Prajna07,Prajna04}. An example of barrier certificate is the stability, which ensures the convergence of states to an equilibrium point and consequently that there exists a region from where any trajectory remains inside this region. The stability can be ensured based on Lyapunov theory (for a wide class of model abstractions)~\cite{tabuada2009verification}, or linear system stability theory for discrete simulations~\cite{Bessa17}.

Generally, properties that must hold for ECPSs are related to safety/reachability and liveness, and the earliest studies on HS and ECPS proposed approaches based on computational tree logic and linear temporal logic~\cite{Alur93}, in order to specify properties and verify those systems. In the last decade, the signal temporal logic~\cite{Maler04,Donze12} was proposed and have been used since then for this purpose, thus providing more suitable operators for dynamic systems and signal processing. Therefore, model-checking procedures have been employed to verify those properties, even under undecidability conditions via approximations, {\it e.g.}, bounding the number of steps from initial states, which leads to bounded model checking (BMC)~\cite{Veanes09}.

@@

Finally, it is worth mentioning that there are other alternatives for ECPS modeling and analysis, which are not based on verification, but instead on simulators~\cite{Phan2014,Nakajima12,Junjie12}. Although such approaches generally fail in ensuring that some properties hold, they present good scalability, which is ideal for large scale and complex systems, which are hardly modeled through automata- or logig-based approaches. Modelica~\cite{Junjie12} and Simulink/MATLAB~\cite{Nakajima12} are important examples of widely used simulators for ECPSs. In order to achieve a balanced trade-off between ECPS-simulation scalability and correctness of verification methods, statistical model checking have been employed to formally verify ECPSs~\cite{Clarke11}.

In addition, Shoukry {\it et al.} \cite{Shoukry} tackled another important problem related to ECPSs: security. Indeed, the essence of ECPSs claims for sophisticated interfacing ({\it e.g.}, sensors and network connections), which normally leads to security issues~\cite{cpsattack}, given that wrong control actions can be generated through corrupted measurements. The mentioned authors proposed a methodology for estimating physical-system state, through formal methods, even if sensor inputs are contaminated with erroneous or intentionally corrupted data, in order to still be able to execute suitable control commands. Cyber-security is also a concern of Fiore {\it et al.}, who aim to ensure secure-state estimation of UAV systems under sparse attacks~\cite{Fiore17}. Indeed, many situations can be regarded as satisfiability problems and solved, for instance, with approaches based on Satisfiability Modulo Theories (SMT)~\cite{Araujo16,TrindadeDAES2016,Rahman13}, which indicate many links with other research areas and might lead to a unified approach regarding system verification, security enforcement, and interaction handling.

%--------------------------------------
\subsection{Control Synthesis of Cyber-physical Systems}
\label{ssec:synthesis}
%--------------------------------------
